\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2016-07-22 redaction/dhp_subETSG.tex
Version sans gras soumise ETSG

Table \ref{tab:summary_results} yields the following comments. First, the magnitude of overall transport costs is sizeable, either considering the average value over the period or even in the most recent years. Over 1974-2013, they correspond to a mean increase of the export price by 5.8\% for ocean shipping, and by 5.1\% for air shipping. \textbf{comparer avec la litterature?}. Second, when decomposing transport costs into an additive and a multiplicative component, we find that the per-unit cost dimension is sizeable. Over the whole period, additive costs are 2.85\% of fob unit values for ocean shipping  - and ad-valorem ones being equal to 3.22\%. These values are respectively equal to 1.9\% and 2.5\% for air transport. For both transport modes, the omission of the additive term seriously biases the multiplicative term upward. In quantitative terms, the omission of the additive term leads to overestimate the iceberg component by roughly a factor 2.

@@

\begin{figure}[htbp]
\caption{Decomposing Transport costs (3 digits}\label{fig:decomp_TC_3d}
\begin{center}
\begin{tabular}{cc}
{\small (a) Air } & {\small (b) Vessel}\\
\includegraphics[width=3.5in, height=3in]{Fig2a_decompTC_air_3d.pdf}
& \includegraphics[width=3.5in,height=3in]{Fig2b_decompTC_vessel_3d.pdf} \\
\end{tabular}
\end{center}
\end{figure}

@@

In this section, we explore the performances of each type of model (with and without additive costs) in fitting the observed cif-fob prices gap, in order to deliver a more systematic diagnosis about the importance of additive costs. To do so, we rely on several standard measures of fit. The first indicator is through comparing $R^{2}$. However, its use is far from being straightforward when evaluating non-linear estimates. \textbf{mettre une reference sur le point de la footnote}\footnote{$R^2$ is based on the underlying assumption that the adjusted model is a linear one. In a non-linear context, R-squared is therefore inappropriate, strictly speaking. However, if the error distribution is approximately normal, a standard metric like R-squared remains informative on the quality of adjustment.}. As a complement, we provide the Standard Error of Regression (SER), which represents the average distance that the observed values fall from the regression line. Smaller values are better because it indicates that the observations are closer to the fitted line. We also report the log-likelihood function, and two measures derived, the Akaike Information Criterion (AIC) and the log-likelihood (LL) ratio test. A decrease in the log-likelihood function points to a better quality-of-fit. However, the likelihood function systematically decreases with the number of parameters included; the AIC criterion allows for correcting this overfitting by including a penalty in the computation of the statistic, so that AIC stat $= 2 \times \textrm{number of parameters} - 2 \times \textrm{Likelihood} $. Once again, the preferred model is the one with the minimum AIC value. Finally, the log-likelihood ratio test statistic compares systematically the likelihood of the Unrestricted model (\emph{UR}, including an additive term, i.e. Equation (\ref{eq:model_with_add}) and the Restricted one (\emph{R}, i.e. Equation (\ref{eq:model_nlI})). The null tested is that the two models are statistically equivalent.\medskip

@@

Going deeper into the examination of the levels of these goodness-of-fit statistics reveals interesting differences. First, multiplicative costs have more explanative power in the Vessel sector, where they account for 39\% of the variance of the CIF/FOB ratio, versus 31\% for the Air sector (considering the mean values of the $R^2$ across transport modes). As well, the standard error of the regression based only on the iceberg term is consistently lower for Vessel than for Air. \textbf{et pourtant, la part des additifs dans vessel est plus elevée, 48.2\% en moyenne vs 42.3\% pour AIr, cf le .xls. A dire? Faire reference a la figure 2.}

Second, the dynamics over time also reveals that the quality of fit decreases strongly after 2000, for both transport modes. For Air, the additive term explains less and less variance over time: In 2000, the inclusion of the additive term allows adding more than 30\% of explanatory power, and decreases the standard error of the regression by 14 percentage points;  in 2010 and 2013, it is hardly more than 10\% for the $R^{2}$, and 7 percentage points for the SER. The picture is quite different for Vessel. In this case, the inclusion of the additive term increases the R$^{2}$ by a percentage very similar whatever the considered year (between 12 and 18\%); a similar conclusion can be drawn for the SER, which decreases by 8-11 percentage points across the years when additive costs are included. Rather, the deteriorating performance of the model in terms of goodness-of-fit from 2000 mainly comes from the decreasing performance of the iceberg term. \textbf{so what? pas tres punchy}\smallskip

>>>>

Table \ref{tab:summary_results} yields the following comments. First, the magnitude of overall transport costs is sizeable, either considering the average value over the period or even in the most recent years. Over 1974-2013, they correspond to a mean increase of the export price by 5.8\% for ocean shipping, and by 5.1\% for air shipping. Second, when decomposing transport costs into an additive and a multiplicative component, we find that the per-unit cost dimension is sizeable. Over the whole period, additive costs are 2.85\% of fob unit values for ocean shipping  - and ad-valorem ones being equal to 3.22\%. These values are respectively equal to 1.9\% and 2.5\% for air transport. For both transport modes, the omission of the additive term seriously biases the multiplicative term upward. In quantitative terms, the omission of the additive term leads to overestimate the iceberg component by roughly a factor 2.

@@

\begin{figure}[htbp]
\caption{Decomposing Transport costs (3 digits}\label{fig:decomp_TC_3d}
\begin{center}
\begin{tabular}{cc}
{\small (a) Air } & {\small (b) Vessel}\\
\includegraphics[width=3in, height=2.5in]{Fig2a_decompTC_air_3d.pdf}
& \includegraphics[width=3in,height=2.5in]{Fig2b_decompTC_vessel_3d.pdf} \\
\end{tabular}
\end{center}
\end{figure}

@@

In this section, we explore the performances of each type of model (with and without additive costs) in fitting the observed cif-fob prices gap, in order to deliver a more systematic diagnosis about the importance of additive costs. To do so, we rely on several standard measures of fit. The first indicator is through comparing $R^{2}$. However, its use is far from being straightforward when evaluating non-linear estimates.\footnote{$R^2$ is based on the underlying assumption that the adjusted model is a linear one. In a non-linear context, R-squared is therefore inappropriate, strictly speaking. However, if the error distribution is approximately normal, a standard metric like R-squared remains informative on the quality of adjustment.}. As a complement, we provide the Standard Error of Regression (SER), which represents the average distance that the observed values fall from the regression line. Smaller values are better because it indicates that the observations are closer to the fitted line. We also report the log-likelihood function, and two measures derived, the Akaike Information Criterion (AIC) and the log-likelihood (LL) ratio test. A decrease in the log-likelihood function points to a better quality-of-fit. However, the likelihood function systematically decreases with the number of parameters included; the AIC criterion allows for correcting this overfitting by including a penalty in the computation of the statistic, so that AIC stat $= 2 \times \textrm{number of parameters} - 2 \times \textrm{Likelihood} $. Once again, the preferred model is the one with the minimum AIC value. Finally, the log-likelihood ratio test statistic compares systematically the likelihood of the Unrestricted model (\emph{UR}, including an additive term, i.e. Equation (\ref{eq:model_with_add}) and the Restricted one (\emph{R}, i.e. Equation (\ref{eq:model_nlI})). The null tested is that the two models are statistically equivalent.\medskip

@@

Going deeper into the examination of the levels of these goodness-of-fit statistics reveals interesting differences. First, multiplicative costs have more explanative power in the Vessel sector, where they account for 39\% of the variance of the CIF/FOB ratio, versus 31\% for the Air sector (considering the mean values of the $R^2$ across transport modes). As well, the standard error of the regression based only on the iceberg term is consistently lower for Vessel than for Air.

Second, the dynamics over time also reveals that the quality of fit decreases strongly after 2000, for both transport modes. For Air, the additive term explains less and less variance over time: In 2000, the inclusion of the additive term allows adding more than 30\% of explanatory power, and decreases the standard error of the regression by 14 percentage points;  in 2010 and 2013, it is hardly more than 10\% for the $R^{2}$, and 7 percentage points for the SER. The picture is quite different for Vessel. In this case, the inclusion of the additive term increases the R$^{2}$ by a percentage very similar whatever the considered year (between 12 and 18\%); a similar conclusion can be drawn for the SER, which decreases by 8-11 percentage points across the years when additive costs are included. Rather, the deteriorating performance of the model in terms of goodness-of-fit from 2000 mainly comes from the decreasing performance of the iceberg term. \smallskip
